---
title: "Assignment 1 - Stats R programming"
author: "Thrisha Rajkumar"
date: "`r Sys.Date()`"
output: html_document
---

1. There is a block of YAML header at the start of the document which gives key information
eg. title and output format.
2. You can create headings of sections with “#”. Similarly, headings of subsections can
be created with “##” and so on.
3. You can embed blocks of R code by using ``` {r} before the R code and ``` afterwards.
4. By default both the code and its output are displayed. If you only want to include the
output but not the code itself include the option echo = FALSE in the code prefix. If you
don’t want to include either then include the option include = FALSE.
5. You can include hyperlinks by writing <url-name>.

Step 1. Insert a block of R code. In this block of code, create a data frame animals df,
using the same code you did in Question 1. In R Markdown, a block of code starts with ```
{r} and ends with ```
Step 2. Insert another block of code to print animals df.
Step 3. You can then generate an HTML file with the “Knit” button just above your
scripting window. Check what you have in the HTML file.


```{r}
animals <- c("Snake", "Ostrich", "Cat", "Spider")

num_legs <- c(0,2,4,8)

#combine the vectors in a dataframe calld animals_df
#which has two columns - names of animals and their 
#respective number of legs 

#generating a dataframe
animals_df <- data.frame(animals, num_legs)


ls()
```

```{r}
animals_df

```
