---
title: "Assignment 3"
author: "Thrisha Rajkumar"
date: "`r Sys.Date()`"
output: html_document
---


## Load Packages

We need to load two packages, namely “Stat2Data” and “tidyverse”, before
answering the questions. If they haven’t been installed on your computer, please use
“install.packages()” to install them first.
1. Load the tidyverse package:
2. Load the Stat2Data package and then the dataset Hawks:

```{r}
library(tidyverse)
library(Stat2Data)
data("Hawks")
```

## Visualization

 using ggplot2
 
## (Q1) After the dataset Stat2Data was loaded sucesfully, create a smaller dataset with the code below

```{r}
hawksSmall<-
drop_na(select(Hawks,Age,Day,Month,Year,CaptureTime,Species,Wing,Weight,Tail))
dim(hawksSmall)
head(hawksSmall)
```

## (Q2) Use a combination of the functions “ggplot()” and “geom_histogram” to create a histogram plot of the weights of Hawks within the hawksSmall data frame with bin widths of 10. Your result should look something like the following.

```{r}
ggplot(data=hawksSmall,aes(x=Weight))+geom_histogram(binwidth=10)+
xlab("Weight")+ylab("Count")

```

## (Q3) Similar to (Q2), use the “geom_density()” function to create two density plots of the tail lengths of Hawks within the hawksSmall data frame, with parameters “adjust=0.5” and “adjust=2”, respectively. Your results should look like this:


```{r}
ggplot(data = hawksSmall, aes(x=Tail))+geom_density(adjust=0.5)+xlab("Tail Length")+ylab("Count")
ggplot(data = hawksSmall, aes(x=Tail))+geom_density(adjust=2)+xlab("Tail Length")+ylab("Count")
```

Can you explain the difference between the two plots? How many modes does the distribution of Hawk tail lengths have in each of the plots.

The second plot has a curve that is smoother version of the first plot. This is because the adjust argument in the geom_density() function adjusts the bandwidth of the density plot.A larger "adjust" corressponds to a larger bandwidth, hencce a smoother function in the plot, whilst a smaller "adjust" create a less smooth density plot which is mor esensitive to the local behavior of the data.

Three modes are visible in the first plot. Two modes are visible in the second plot. 

##(Q4) Based on the answer from (Q3), can you create the following plot?


```{r}
ggplot(data=hawksSmall, aes(x=Tail, color=Species))+geom_density(adjust=2)+xlab("Tail length")+ylab("Density")
```


## (Q5) Violin plot: Use the ggplot and geom_violin() functions to create the following violin plot for the three species. 

```{r}
ggplot(data=hawksSmall, aes(x=Tail, y=Species, fill=Species))+geom_violin()+xlab("Tail length")+ylab("Density") +theme_bw()
```


## (Q6) Scatter plot
## Generate a plot similar to the following plot using the ggplot() and geom_point()functions.
## 1. How many aesthetics are present within the following plot?
There are four aesthetics: 
(1) The tail length is mapped to the horizontal position, 
(2) The weight is mapped to the vertical position, 
(3) The species is mapped to the colour and 
(4) The species is mapped to shape. 

## 2. What are the glyphs within this plot?
The glyphs are the small elements (in this particular case represented by the shapes) corresponding to individual cases

## 3. What are the visual cues being used within this plot? 
The visual cues include horizontal and vertical position, shape, color. 

```{r}
ggplot(hawksSmall, aes(x=Tail, y=Weight, color=Species, shape=Species)) + geom_point() + xlab('Tail (mm)') + ylab('Weight (mm)') + theme_bw()
```


## (Q7) Trend lines and facet wraps:
##Generate the following plot using the ggplot(), geom_point(), geom_smooth() and facet_wrap() functions. Note that in the facet plot, the three panels use different scales.

## 1. What are the visual cues being used within this plot?
The visual cues include horizontal and vertical position, shape, color and direction. 

## 2. Based on the plot below, what can we say about the relationship between the weight of the hawks and their tail lengths? 
Based on this sample, longer tail lengths appear to be predictive of larger weights within each species. 

```{r}
ggplot(hawksSmall, aes(x=Tail, y=Weight, color=Species)) +
  geom_point() + facet_wrap(~Species, scales = 'free') + geom_smooth(method = 'lm') + xlab('Tail (mm)')+ylab('Weight (mm)') +theme_bw()
```

## (Q8) Adding annotations
## First, compute the “Weight” and the “Tail” of the heaviest hawk in the dataset. You can use filter() and select() function to select proper data. 

```{r}
max_weight<-max(hawksSmall$Weight, na.rm=TRUE)
max_weight_row = head(hawksSmall %>% filter(Weight>=max_weight), 1)
max_weight_tail = max_weight_row$Tail
```
Second, reuse the code that you create from Q(3), adding an arrow and an annotation to indicate the heaviest hawk. Your result should look similar to this: 

```{r}
ggplot(hawksSmall, aes(x=Tail, y=Weight, color=Species)) + geom_point() + xlab('Tail (mm)') + ylab('Weight (mm)') + theme_bw() + geom_curve(x=max_weight_tail,xend=max_weight_tail,y=max_weight-200,yend=max_weight, arrow=arrow(length=unit(0.1,"cm")), color="black")+ geom_text(x=max_weight_tail,y=max_weight-200,label='heaviest hawk',size=3,color="black")

```


## 2. Finite probability spaces
## Now, let’s move on to consider probability on finite sample spaces, which is covered in Lecture 8.
## 2.1 Sampling with replacement 


for positive integers n and k, 
\[
\binom{n}{k} =\frac{n!}{k!(n-k)!}
\]

gives the number of subsets of size k from a set of n objects.

You can compute this number straightforwardly within “R” via the choose function
“choose()”. For example, if we want to compute the number of different subsets of
size 3 from a collection of size 8 we would compute: 

```{r}
choose(8,3)
```

Suppose we have a bag containing 10 spheres. This includes 3 red spheres and 7 blue spheres. 

Let’s suppose that we draw a sphere at random from the bag (all spheres have equal
probability of being drawn). We record its colour and then return the sphere to the
bag. This process is repeated 22 times. This is an example of sampling with
replacement since the spheres are replaced after each draw. 



(Q1) Write down a mathematical expression for the probability that z out of the 22 selections were red spheres (here z ∈ {0,1, ⋯ ,22}). 


Let A be the event that z out of the 22 selections were red spheres. The probability is

\[
P(A) = \binom{22}{z} \left( \frac{3}{10}^{10}\right) \left(\frac{7}{10}^{22-x}\right)
\]

## (Q2) Next write an R function called “prob_red_spheres()” which takes z as an argument and computes the probability that 𝑧 out of a total of the 22 balls selected are red. 

```{r}
num_red_balls<-3
num_blue_balls<-7
total_draws<-22
prob_red_spheres<-function(z){
 total_balls<-num_red_balls+num_blue_balls
 log_prob<-log(choose(total_draws,z))+
 z*log(num_red_balls/total_balls)+(total_drawsz)*log(num_blue_balls/total_balls)
 return(exp(log_prob))
 prob_red_spheres(10)
}

```

## (Q3) Generate a data frame called “prob_by_num_reds” with two columns “num_reds” and “prob”. The “num_reds” column should contain numbers 1 through 22 and the “prob” column should give the associated probability of selecting that many reds out of a total number of 22 selections. 

```{r}
prob_by_num_reds <- data.frame(num_reds=seq(22)) %>%
 mutate(prob=prob_red_spheres(num_reds))
prob_by_num_reds %>% head(3)
```






















